---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# biology

An R package to faciliate the fitting and exploration of growth models to invertebrate tagging data. In particular there is emphasis on the inverse logistic curve, but also functions relating to the dose-response and Michaelis-Menton curves.

To install this development version from [GitHub](https://www.github.com/) you can use:

```{r echo=TRUE, eval=FALSE}
if (!require(devtools)){install.packages("devtools")} 

devtools::install_github("https://github.com/haddonm/biology")
```

There are currently no vignettes, although each function has its own help page.

If you were to type _methods("plot")_ into the console you will find both plot.bootIL and plot.IL in the list. There are help pages for those, as well as for print.IL and summary.IL. To see the contents of the objects output by, say, _fitIL_, you can use, for example, _str(ans)_. 

Some details of the package can be found using _packageDescription("biology")_.

* 28/06/2020 Initial combination of the older packages invLogistic and maturity.

* 04/07/2020 added extra functionality for plots and analyses. 

* 12/07/2020 0.0.0.900 Modified plot.IL and plotmodelIL. Amended fitDR, but still have to develop code to use the dose-response curve. No Errors, Warnings, or Notes

* 07/08/2020 0.0.0.850 developed the use of alternative dose response curve, using two alternative functions for the standard deviation. The first was constant and the second was an inverse relationship with initial length.

* 09/08/2020 0.0.0.800 Implemented a general funciton for fitting growth curves to taggign growth data. It needs a function to define which growth fucntion should be fitted, and a function to describe how the variance around the predicted growth increments should change as the predicted growth increments change.


Malcolm Haddon

Hobart, June 28, 2020

##
Haddon, M., Mundy, C., and D. Tarbath (2008) Using an inverse-logistic 
    model to describe growth increments of blacklip abalone (_Haliotis rubra_) 
    in Tasmania. _Fishery Bulletin_ __106__:58-71

